---
title: "Grade analysis"
author: "Joséphine Laguardia"
format: html
---

```{r configuration}
#| message: false
here::i_am("r-101-grade-analysis-fork.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(vroom)
```

# Question 1

```{r question 1}
#| message: false
grades <- vroom("grades.csv")
```

# Question 2

```{r}
knitr::kable(grades |> 
               summarise(min(Exam, na.rm = TRUE), 
                         max(Exam,na.rm = TRUE), 
                         median(Exam,na.rm = TRUE), 
                         mean(Exam,na.rm = TRUE)),
             col.names = c("Minimum grade", "Maximum grade", "Median grade", "Mean grade")
)
```

# Question 3

```{r}
miss_exam <- grades |> 
  count(is.na(Exam))
```

`r miss_exam[2,2]` students didn't take the final exam.

# Question 4

```{r}
ggplot(grades, aes(x=Exam)) +
  geom_bar(fill = "darkblue",na.rm=TRUE)
```

# Question 5

```{r}
knitr::kable(grades |> 
  group_by(Group) |>
  summarise(n()),
  col.names = c("Group", "Number of students"))
```

# Question 6

```{r}
gr <- ggplot(grades, aes(x=Group))+
  geom_bar(fill = "darkviolet" ,na.rm=TRUE)
gr + coord_flip()
```

# Question 7

First representation:

```{r}
ggplot(grades, aes(x=Exam))+
  geom_bar(na.rm=TRUE)+
  facet_wrap(~Group)
```

Second representation:

```{r}
ggplot(grades, aes(x=Group, fill=Exam))+
  geom_bar(na.rm=TRUE)
```


# Question 8

```{r}
missed_exam_group <- grades |>
  group_by(Group) |>
  summarise(miss_exam_group= sum(is.na(Exam)))
knitr::kable(missed_exam_group,
             col.names= c("Group", "Number of students who missed the exam"))
```

# Question 9

```{r}
ggplot(missed_exam_group, aes(miss_exam_group, Group))+
  geom_col(fill = "darkred")+
  scale_x_continuous("Number of students who missed the exam")
```

# Question 10

```{r}
long_grades <- grades |>
  pivot_longer(cols = -c('Id','Group'), names_to = "name", values_to = "value")
```
# Question 11

The number of missing grades in total for each student is:

```{r}
missing_grades <- long_grades |>
  group_by(Id) |>
  summarise(number_missing_grades = sum(is.na(value)))
knitr::kable(missing_grades,
             col.names = c("Student Id", "Number of missing grades"))

```
# Question 12

Here is the graphical representation of the distribution of the missing grades per students:

```{r}
ggplot(missing_grades, aes(Id,number_missing_grades))+
  geom_col()+
  scale_y_continuous("Number of missing grades")
```

# Question 13

```{r}
library(stringr)
```

```{r}
knitr::kable(long_grades |>
               group_by(Group)|>
               filter(str_starts(name, "Exam")) |>
               summarise(sum(is.na(value))),
             col.names = c("Group", "Number of students who missed the Exam"))
```
# Question 14

```{r}
missing_online <- long_grades |>
  group_by(Id) |>
  filter(str_starts(name, "Online")) |>
  summarise(number_missing_online = sum(is.na(value)))
knitr::kable(missing_online,
             col.names = c("Id", "Number of missing grades for the online tests"))
             
```

# Question 15

```{r}
# first we merge the two data frames : missing_online and grades
new_grades <- merge(grades, missing_online, by= "Id")
# then we represent the conditional distribution
ggplot(new_grades, aes(x= number_missing_online, y = Exam))+
  geom_point(na.rm=TRUE) +
  scale_x_continuous("Number of missing grades for the online tests")
```
# Question 16

```{r}
mcq_missing <- long_grades |>
  group_by(Id) |>
  filter(str_starts(name, "MCQ")) |>
  summarise(mcq_missing_true = anyNA(value))
knitr::kable(mcq_missing,
             col.names = c("Id", "Missed"))
```
# Question 17

```{r}
# first, we merge grades and mcq_missing
grades2 <- merge(grades, mcq_missing, by = "Id")
# then we create the table with the percentage
percentage_mcq <- grades2 |>
  group_by(Group) |>
  summarise(P_missed = 100*sum(mcq_missing_true)/n())
knitr::kable(percentage_mcq,
             col.names = c("Group","P_missed"))
```
# Question 18

```{r}
# first, we compute the average grade at the Exam per group
average_grade <- grades |>
  group_by(Group) |>
  summarise(average = sum(Exam,na.rm=TRUE)/n())
# then, we merge average_grade and percentage_mcq
merged <- merge(average_grade, percentage_mcq, by="Group")
# then we represent the average grade as a function of P_missed
ggplot(merged, aes(x= P_missed, y= average))+
  geom_point()
```

